---
title: "Data_Cleaning_141"
output: pdf_document
date: "2024-05-15"
---
### Loading the data
```{r message = FALSE, warning = FALSE}
library(readr)
library(dplyr)
library(stringr)
old_pre <- read_csv("old_pre.csv")
old_post <- read_csv("old_post.csv")
new_pre <- read_csv("new_pre.csv")
new_post <- read_csv("new_post.csv")
```

### Removing inessential columns such as timestamp, date of survey, last name, minor, extra elaboration on gender identity, and name of the community partner organization
```{r}
old_pre <- old_pre[,-c(1, 2, 4, 7, 9)]
old_post <- old_post[, -c(1, 2, 4, 7, 9)]
new_pre <- new_pre[,-c(1, 2, 6, 8)]
new_post <- new_post[, -c(1, 2, 5, 7, 9)]
```

### Renaming Columns
```{r}
names(old_pre) <- c("ID", "Course", "Major", "Gender", "Ethnicity", "First_Gen", "Mother_Edu", "Father_Edu", "Transfer", "GPA", "Academic_Confidence", "Sense_Belonging", "Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Text1", "Text2")
names(old_post) <- c("ID", "Course", "Major", "Gender", "Ethnicity", "First_Gen", "Mother_Edu", "Father_Edu", "Transfer", "GPA", "Academic_Confidence", "Sense_Belonging", "Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Text1", "Text2")
names(new_pre) <- c("ID", "Course", "Major", "Gender", "Ethnicity", "First_Gen", "Mother_Edu", "Father_Edu", "Transfer", "GPA", "Academic_Confidence", "Sense_Belonging", "Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Text1", "Text2")
names(new_post) <- c("ID", "Course", "Major", "Gender", "Ethnicity", "First_Gen", "Mother_Edu", "Father_Edu", "Transfer", "GPA", "Academic_Confidence", "Sense_Belonging", "Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Text1", "Text2")
```

### Removing the observations with duplicate ID numbers
```{r}
duplicate_ids_old_pre <- old_pre$ID[duplicated(old_pre$ID) | duplicated(old_pre$ID, fromLast = TRUE)]
duplicate_ids_old_post <- old_post$ID[duplicated(old_post$ID) | duplicated(old_post$ID, fromLast = TRUE)]
duplicate_ids_new_pre <- new_pre$ID[duplicated(new_pre$ID) | duplicated(new_pre$ID, fromLast = TRUE)]
duplicate_ids_new_post <- new_post$ID[duplicated(new_post$ID) | duplicated(new_post$ID, fromLast = TRUE)]

old_pre <- old_pre[!old_pre$ID %in% duplicate_ids_old_pre, ]
old_post <- old_post[!old_post$ID %in% duplicate_ids_old_post, ]
new_pre <- new_pre[!new_pre$ID %in% duplicate_ids_new_pre, ]
new_post <- new_post[!new_post$ID %in% duplicate_ids_new_post, ]
```

### Getting rid of the observations whose IDs are NA and observations whose IDs aren't in both the pre_data and the post_data
```{r}
# old data
old_pre <- old_pre[!is.na(old_pre$ID), ]
old_post <- old_post[!is.na(old_post$ID), ]

ID_old_post <- c()
ID_old_pre <- c()
for(i in 1:length(old_pre$ID)){
  for(j in 1:length(old_post$ID)){
    if(old_pre$ID[i] == old_post$ID[j]){
      ID_old_post <- c(ID_old_post, j)
      ID_old_pre <- c(ID_old_pre, i)
    }
  }
}

# new data
new_pre <- new_pre[!is.na(new_pre$ID), ]
new_post <- new_post[!is.na(new_post$ID), ]

ID_new_post <- c()
ID_new_pre <- c()
for(i in 1:length(new_pre$ID)){
  for(j in 1:length(new_post$ID)){
    if(new_pre$ID[i] == new_post$ID[j]){
      ID_new_post <- c(ID_new_post, j)
      ID_new_pre <- c(ID_new_pre, i)
    }
  }
}

# keeping the ones that have matching IDs in both pre & post
old_post <- old_post[ID_old_post,]
old_pre <- old_pre[ID_old_pre,]
new_post <- new_post[ID_new_post,]
new_pre <- new_pre[ID_new_pre,]
new_pre <- new_pre[new_pre$ID != 2794, ]
new_post <- new_post[new_post$ID != 2794, ]
```


### Replacing the NA values in the questions with ranking responses with the median values for each of them
```{r}
# old data
columns_with_ranking <- c("Academic_Confidence", "Sense_Belonging", paste0("Q", 1:15))
old_post[columns_with_ranking] <- lapply(old_post[columns_with_ranking], function(x) {
  x[is.na(x)] <- median(x, na.rm = TRUE)
  return(x)
})

old_pre[columns_with_ranking] <- lapply(old_pre[columns_with_ranking], function(x) {
  x[is.na(x)] <- median(x, na.rm = TRUE)
  return(x)
})

# new data
new_post[columns_with_ranking] <- lapply(new_post[columns_with_ranking], function(x) {
  x[is.na(x)] <- median(x, na.rm = TRUE)
  return(x)
})

new_pre[columns_with_ranking] <- lapply(new_pre[columns_with_ranking], function(x) {
  x[is.na(x)] <- median(x, na.rm = TRUE)
  return(x)
})
```

### Cleaning up the "Course" variable
```{r}
# old data
replacements <- list(
  "CCAS100XP" = c("23S-CCAS 100XP", "CCAS 100XP", "CCAS 100xp", "ccas100xp", "CCAS100XP, CCAS143, CCASM173, PSYCH137C", "CCAS100XP"),
  "CESC50XP" = c("CECS50XP", "CESC 50XP", "CSEC50XP", "Community Engagement and Social Change, CESC50XP", "CESC50XP, PhySci111B, MCDB60, PhySci199, LS192B PhySci189, LS", "Center for Community Engagement, CESC50XP", "23S-CESC-50XP-LEC-1", "CECS50XP", "CESC50XP"),
  "CESC172XP" = c("CESC 172XP", "CSEC 172xp", "PHYSCI5, ANTHRO2, PHYSICS5B, CESC172XP", "CESC 172XP, CESC 191BX, Neuro M116A", "CESC172XP"),
  "EDUC144XP" = c("EDUC 144XP", "Educ144XP", "EDUCATION 144XP: PEDAGOGIES OF GLOBAL CITIZENSHIP EDUCATION", "EDUC144XP EDUC130AX INFSTD180", "EDUC144XP", "PHYSCI5, ANTHRO2, PHYSICS5B, CESC172XP")
)
for (new_value in names(replacements)) {
  old_pre$Course[old_pre$Course %in% replacements[[new_value]]] <- new_value
}


replacements_2 <- list(
  "CCAS100XP" = c("23S-CCAS-100XP", "23S-CCAS100XP", "CCAS 100xp", "CCAS 100XP", "CCAS100XP", "CCAS100XPSEM1"),
  "CESC50XP" = c("23SCESC50XP", "cesc 50xp", "CESC 50XP", "CESC-50XP", "CESC50XO=P", "CESC50xp", "CESC50XP"),
  "CESC172XP" = c("CESC 172 xp", "CESC 172XP", "CESC-172XP", "CESC172XP", "CESC172XR"),
  "EDUC144XP" = c("144 XP", "EDUC 144XP", "EDUC144 XP", "Educ144XP", "EDUC144XP")
)
for (new_value in names(replacements_2)) {
  old_post$Course[old_post$Course %in% replacements_2[[new_value]]] <- new_value
}

# new data
new_pre <- new_pre %>% filter(Course != "ED130AX")
replacements_3 <- list(
  "CESC191AX" = "CESC191AX",
  "ENGCOMP3DX" = c("ENGCOMP3DX", "CESC3DX"),
  "CESC50XP" = "CESC50XP",
  "DISSTD138XP" = "DISSTD138XP",
  "DISSTDM172XP" = "DISSTDM172XP",
  "ECONM188XP" = "ECONM188XP"
  )
for (new_value in names(replacements_3)) {
  new_pre$Course[new_pre$Course %in% replacements_3[[new_value]]] <- new_value
}

replacements_4 <- list(
  "ENGCOMP3DX" = "ENGCOMP3DX",
  "CESC50XP" = c("CESC 50XP", "CESC50xp", "CESC50XP"),
  "DISSTD138XP" = "DISSTD138XP",
  "DISSTDM172XP" = c("DISSTDM172XP", "DISSTD M172XP"),
  "ECONM188XP" = "ECONM188XP"
  )
for (new_value in names(replacements_4)) {
  new_post$Course[new_post$Course %in% replacements_4[[new_value]]] <- new_value
}
```

### Cleaning up the "Ethnicity" variable
```{r}
# old data
replacements_5 <- list(
  "Asian / Pacific Islander" = "Asian / Pacific Islander",
  "Hispanic/Latinx" = "Hispanic/Latinx",
  "White" = "White",
  "Black or African American" = "Black or African American",
  "Multiple Ethnicity / Other" = c("American Indian or Alaskan Native, Hispanic/Latinx", "Asian / Pacific Islander, Black or African American, Hispanic/Latinx", "Asian / Pacific Islander, Hispanic/Latinx, White", "Asian / Pacific Islander, White", "Asian / Pacific Islander, White, Multiple Ethnicity / Other", "Black or African American, Hispanic/Latinx", "Black or African American, White, Multiple Ethnicity / Other", "Hispanic/Latinx, White", "Hispanic/Latinx, White, Multiple Ethnicity / Other", "Multiple Ethnicity / Other", "White, Multiple Ethnicity / Other")
  
)
for (new_value in names(replacements_5)) {
  old_pre$Ethnicity[old_pre$Ethnicity %in% replacements_5[[new_value]]] <- new_value
}

replacements_6 <- list(
  "Asian / Pacific Islander" = "Asian / Pacific Islander",
  "Hispanic/Latinx" = "Hispanic/Latinx",
  "White" = "White",
  "Black or African American" = "Black or African American",
  "Multiple Ethnicity / Other" = c("American Indian or Alaskan Native;Hispanic/Latinx", "American Indian or Alaskan Native;Hispanic/Latinx;Multiple Ethnicity / Other", "Asian / Pacific Islander;Black or African American;Hispanic/Latinx", "Asian / Pacific Islander;Hispanic/Latinx;White", "Asian / Pacific Islander;White", "Asian / Pacific Islander;White;Multiple Ethnicity / Other", "Black or African American;Hispanic/Latinx", "Black or African American;White", "Black or African American;White;Multiple Ethnicity / Other", "Hispanic/Latinx;White", "Hispanic/Latinx;White;Multiple Ethnicity / Other", "Multiple Ethnicity / Other")
  
)
for (new_value in names(replacements_6)) {
  old_post$Ethnicity[old_post$Ethnicity %in% replacements_6[[new_value]]] <- new_value
}

# new data
replacements_7 <- list(
  "Asian / Pacific Islander" = "Asian / Pacific Islander",
  "Hispanic/Latinx" = "Hispanic/Latinx",
  "White" = "White",
  "Black or African American" = "Black or African American",
  "Multiple Ethnicity / Other" = c("Asian / Pacific Islander, Black or African American", "Asian / Pacific Islander, Black or African American, White, Multiple Ethnicity / Other", "Asian / Pacific Islander, Hispanic/Latinx", "Asian / Pacific Islander, Hispanic/Latinx, White, Multiple Ethnicity / Other", "Asian / Pacific Islander, White", "Asian / Pacific Islander, White, Multiple Ethnicity / Other", "Black or African American, Hispanic/Latinx", "Black or African American, White", "Hispanic/Latinx, White", "Hispanic/Latinx, White, Multiple Ethnicity / Other", "Multiple Ethnicity / Other", "White, Multiple Ethnicity / Other")
  
)
for (new_value in names(replacements_7)) {
  new_pre$Ethnicity[new_pre$Ethnicity %in% replacements_7[[new_value]]] <- new_value
}

replacements_8 <- list(
  "Asian / Pacific Islander" = "Asian / Pacific Islander",
  "Hispanic/Latinx" = "Hispanic/Latinx",
  "White" = "White",
  "Black or African American" = "Black or African American",
  "Multiple Ethnicity / Other" = c("Asian / Pacific Islander, Hispanic/Latinx", "Asian / Pacific Islander, Hispanic/Latinx, White, Multiple Ethnicity / Other", "Asian / Pacific Islander, Multiple Ethnicity / Other", "Asian / Pacific Islander, White", "Asian / Pacific Islander, White, Multiple Ethnicity / Other", "Black or African American, Hispanic/Latinx, Multiple Ethnicity / Other", "Black or African American, White, Multiple Ethnicity / Other", "Hispanic/Latinx, White", "Hispanic/Latinx, White, Multiple Ethnicity / Other", "Multiple Ethnicity / Other", "White, Multiple Ethnicity / Other")
  
)
for (new_value in names(replacements_8)) {
  new_post$Ethnicity[new_post$Ethnicity %in% replacements_8[[new_value]]] <- new_value
}

levels <- c("Asian / Pacific Islander", "Hispanic/Latinx", "White", "Black or African American", "Multiple Ethnicity / Other")

old_pre$Ethnicity <- factor(old_pre$Ethnicity, levels = levels)
old_post$Ethnicity <- factor(old_post$Ethnicity, levels = levels)
new_pre$Ethnicity <- factor(new_pre$Ethnicity, levels = levels)
new_post$Ethnicity <- factor(new_post$Ethnicity, levels = levels)
```

### Categorizing the "Major" variable into "North Campus" and "South Campus"
```{r}
replacements_9 <- list(
  "North Campus" = c(
    "American literature and culture",
    "Anthropology B.S.",
    "Applied Linguistics", 
    "American Literature and Culture",
    "Art",                                               
    "Art History", 
    "Architectural Studies",
    "Business Econ",
    "Business Economics",
    "Business Economic",
    "Business economics",
    "BUSINESS ECONOMICS",
    "Business economics and statistics and data science",
    "Business Economics, Cognitive Science",             
    "Business-economics",                                
    "Business-Economics",
    "Business Economics & Education",                    
    "Business Economics and Psychology",
    "Business Economics (switching to Cognitive Science)",
    "Business Economics; Education and Social Transformation",
    "bus econ",
    "Bus/Econ",
    "CCAS",
    "Chicana & Chicano Studies",
    "Chicana and Chicano Studies",
    "Chicano, Chicana, and Central American Studies",
    "Chicana/o and Central American Studies",
    "chicana/o and central american studies \\ anthropology",
    "Chicana/o Studies",
    "Chicana/o Studies & Political Science",
    "Chicana/o Studies Major",
    "Chicana/o studies and Sociology",
    "Chicano Studies",
    "Chicano Studies & Sociology double major",
    "Chicanx Studies",
    "Chicanx and Central American Studies",
    "China/o Studies",
    "Chicano/a Studies",
    "Chicano/a Studies major",
    "chicano stuides",
    "Chican@ Studies",
    "Communication",
    "communication", 
    "Communications",
    "Comparative Literature",
    "Design Media Arts",
    "DesignMediaArts",
    "Disability Studies and Public Affairs",
    "Econ",
    "ECON",
    "Economics",
    "Economcis",
    "ECONOMICS",
    "Education",
    "Econ, Psych",
    "Environmental Science",                            
    "Fine Art",
    "Disability Studies and Public Affairs",             
    "Ecology, Behavior, and Evolution",
    "EDUCATION AND SOCIAL TRANSFORMATION,",
    "Education & Social Transformation / Public Affairs",
    "Education and social transformation",
    "Education and Social Transformation",
    "Education and Social Transformation and Chicano/a Studies",
    "Education and Social Transformation & Business Economics",
    "Education and Social Transformation, Chicano/a Studies",
    "EDUCATION AND SOCIAL TRANSFORMATION",
    "Education and Social Transformation",
    "Economics and Environmental Science",
    "Economics and Public Affairs",
    "English",
    "Ethnomusicology",
    "English and Chicano Studies",
    "Geography",
    "Geography + Chicano Studies",
    "Geography and Chicano Studies",
    "Global Jazz Studies",
    "global studies",
    "History",
    "Global Studies",
    "Labor Studies",
    "labor studies & chicano/a and central american studies",
    "Labor Studies and Chincao Studies",
    "Labor studies & Chicano studies",
    "Political science",
    "Political Science",
    "Political Science and CCAS",
    "Political Science and Economics",
    "Political Science and Spanish",
    "Political Science, Psychology",
    "pre Business Economic",
    "Political Science and Chicana/o Studies",
    "Political Science and Chicano Studies",
    "Political science and psychology",
    "Political Science, Spanish",
    "Pre-Political Science",
    "Pre-Sociology",
    "psychology",
    "Psych",
    "Psychology",
    "Psychology and Sociology",
    "Psychology but switching to Statistics",
    "psychology",                                        
    "Psychology",                                        
    "Psychology and Disability Studies",                 
    "Public Affairs",                                    
    "Public Affiars" ,                                   
    "sociology",
    "Sociology",
    "Sociology and ccas",
    "Sociology/Spanish",
    "Sociology and CCAS",                                                          
    "Sociology, Psychology",
    "Spanish Community and Culture",
    "Socioloy (currently still pre-business econ but I am switching next quarter)",
    "Theater",
    "Theatre",
    "World Arts and Culture",
    "Chican@ Studies",
    "Chicano, Chicana, and Central American Studies",
    "Chicanx and Central American Studies",
    "Communications",
    "ECON",
    "Economcis",
    "Education",
    "Education and Social Transformation & Business Economics",
    "Education and Social Transformation, Chicano/a Studies",
    "Geography and Chicano Studies",
    "Political Science & Chicano Studies",
    "Political Science",
    "Political Science and Chicana/o Studies",                                     
    "Political Science and Chicano Studies",
    "Political Science and Psychology",
    "Political Science, Spanish: Community and Service",
    "Political Sciene/ CCAS",
    "Political Science, Spanish: Community and Service",
    "Political science",
    "Political Science and Sociology",                   
    "Political Science, Psychology",                     
    "Political Science/Communications",
    "Undeclared - Social Sciences",
    "Pre-Political Science",
    "Pre-Sociology",
    "Pre-Psychology",
    "IDS",                                               
    "Labor Studies",                                     
    "Linguistics and Anthropology", 
    "Public Affairs" ,
    "Public Affairs and Education & Social Transformation",
    "Pre business economics",
    "Pre Political Science",                           
    "Pre-business economics",                          
    "Pre-Economics" ,        
    "Pre Business Economics", 
    "Buisness Economics",
    "Pre- Business Economics",                           
    "Pre-business econ",                                 
    "pre-education",                                     
    "psychology",                                
    "Pre-psychology",                                    
    "Pre-Public Affairs",                                
    "Psych",
    "psych",
    "Sociology, Psychology",
    "Socioloy (currently still pre-business econ but I am switching next quarter)"),

  "South Campus" = c(
    "Aerospace Engineering",
    "Applied Maths & Statistics",
    "Astrophysics",
    "biochemistry",
    "biochem",
    "Biochem",
    "Biochemistry",
    "Bioengineering",
    "Biology",
    "Biophysics",
    "cognitive science",
    "Cognitive science",
    "Cognitive Science",
    "cognitive science & public affairs",                
    "Cognitive Science and Philosophy",
    "Civil Engineering",                                 
    "Cognitive Science",
    "Computer Science",
    "Computational and Systems Biology",                
    "Computational Systems and Biology",
    "Computer Science & Engineering",
    "Computer Science and Engineering",
    "Chemistry",
    "Data Theory",
    "Ecology Behavior and Evolution",
    "Ecology, Behavior, and Evolution",
    "Economics and Statistics & Data Science",
    "Human Biology and Society",
    "Human Biology & Society",
    "Human Biology and Society B.S.",
    "HBS",                                               
    "Human Biology & Society", 
    "human biology and society",
    "Human Biology and Society",                        
    "Human Biology and Society (BS)", 
    "Math/Econ",
    "Mathematics and Economics",
    "Mathematics of Computation"  ,
    "Mathematics/Economics",
    "MCDB",
    "MCDB (Molecular, Cell, Developmental Biology)",
    "Mechanical Engineering",
    "Microbiology, Immunology, and Molecular Genetics",
    "Microbiology, Immunology & Molecular Genetics",
    "MIMG",
    "Molecular cell and developmental biology",
    "Molecular Cell and Developmental Biology (MCDB)",
    "Molecular, Cell and Developmental Biology",
    "Molecular, Cell, and Developmental Biology",
    "Molecular, Cell, and Developmental Biology B.S.",
    "Molecular, Cell, and Developmental Biology",
    "Molecular, Cell, Developmental Biology",
    "Molecular, Cellular and Developmental Biology",
    "Neuroscience",
    "Neuroscience, B.S.",
    "Nursing",
    "Physics",
    "Physics BS",
    "Physiological science",
    "Physiology",
    "Physiological Science",
    "Physiological Sciences",
    "Pre HBS",
    "pre Human Biology and Society",
    "Pre-applied Mathematics",
    "pre-human biology and society (b.s)",
    "Psychobiology",
    "Pre Data Theory",
    "Pre-psychobiology",
    "Pre-psychobiology",
    "Materials science engineering",                     
    "Math and Economics",                                
    "Math of Computation",                               
    "Math/Econ",  
    "Math",
    "Math Econ",
    "Mathematics",                                       
    "Mathematics of Computation",                        
    "Mathematics/Economics",                             
    "MCDB",                                              
    "Mechanical Engineering",                            
    "Microbiology, Immunology and Molecular Genetics" ,  
    "Microbiology, Immunology, and Molecular Genetics" , 
    "Mimg" ,                                             
    "MIMG",                                              
    "Molecular, Cell, and Developmental Biology",        
    "Neuroscience",                                      
    "Neuroscience, B.S",                                 
    "Psychobiology" ,
    "psychobiology" ,
    "Psychobiology & Economics", 
    "Physiological Science" ,                            
    "Physiological Sciences",                            
    "Physiology",                               
    "Statistics",
    "statistics",
    "Statistics and Data Science",
    "Statistics & Data Science",                         
    "statistics and data science",                       
    "Statistics and data science",                       
    "Statistics and Data Science",                       
    "Statistics and Data Science/Economics",
    "Statistics & Data Science and Economics",
    "Statistics and Data Science/Economic",
    "STATS and Data Science",
    "Stats and Date Science",
    "Undeclared - Physical Sciences",
    "Undeclared- Life Sciences",
    "HBS",
    "Undecided life science", 
    "Undecided Life Science",
    "Human Biology & Society",
    "Microbiology, Immunology & Molecular Genetics",
    "Molecular Cell and Developmental Biology (MCDB)",
    "Molecular, Cell, and Developmental Biology B.S.",
    "Molecular, Cellular and Developmental Biology",
    "Physics BS",
    "Statistics and Data Science, Financial Actuarial Mathematics",
    "Undeclared Life Sciences",
    "Undecided life science"),
  "Undecided/Multi-Major/Other" = c("Double Major in Cognitive Science and Philosophy","undeclared", "Undeclared", "cbreynolds17@gmail.com", "Neuroscience, Disability Studies, Individual Major")
)
for (new_value in names(replacements_9)) {
  old_pre$Major[old_pre$Major %in% replacements_9[[new_value]]] <- new_value
}

for (new_value in names(replacements_9)) {
  old_post$Major[old_post$Major %in% replacements_9[[new_value]]] <- new_value
}

for (new_value in names(replacements_9)) {
  new_pre$Major[new_pre$Major %in% replacements_9[[new_value]]] <- new_value
}

for (new_value in names(replacements_9)) {
  new_post$Major[new_post$Major %in% replacements_9[[new_value]]] <- new_value
}
```

### Setting the categorical variables as factors and NA omit
```{r}
# old data
old_pre$Course <- as.factor(old_pre$Course)
old_pre$Major <- as.factor(old_pre$Major)
old_pre$Gender <- as.factor(old_pre$Gender)
old_pre$First_Gen <- as.factor(old_pre$First_Gen)
old_pre$Transfer <- as.factor(old_pre$Transfer)
old_pre$Father_Edu <- as.factor(old_pre$Father_Edu)
old_pre$Mother_Edu <- as.factor(old_pre$Mother_Edu)
old_pre$GPA <- as.factor(old_pre$GPA)

old_post$Course <- as.factor(old_post$Course)
old_post$Major <- as.factor(old_post$Major)
old_post$Gender <- as.factor(old_post$Gender)
old_post$First_Gen <- as.factor(old_post$First_Gen)
old_post$Transfer <- as.factor(old_post$Transfer)
old_post$Father_Edu <- as.factor(old_post$Father_Edu)
old_post$Mother_Edu <- as.factor(old_post$Mother_Edu)
old_post$GPA <- as.factor(old_post$GPA)

# new data
new_pre$Course <- as.factor(new_pre$Course)
new_pre$Major <- as.factor(new_pre$Major)
new_pre$Gender <- as.factor(new_pre$Gender)
new_pre$First_Gen <- as.factor(new_pre$First_Gen)
new_pre$Transfer <- as.factor(new_pre$Transfer)
new_pre$Father_Edu <- as.factor(new_pre$Father_Edu)
new_pre$Mother_Edu <- as.factor(new_pre$Mother_Edu)
new_pre$GPA <- as.factor(new_pre$GPA)

new_post$Course <- as.factor(new_post$Course)
new_post$Major <- as.factor(new_post$Major)
new_post$Gender <- as.factor(new_post$Gender)
new_post$First_Gen <- as.factor(new_post$First_Gen)
new_post$Transfer <- as.factor(new_post$Transfer)
new_post$Father_Edu <- as.factor(new_post$Father_Edu)
new_post$Mother_Edu <- as.factor(new_post$Mother_Edu)
new_post$GPA <- as.factor(new_post$GPA)
```

### Saving the data
```{r}
write.csv(old_pre, file = "cleaned_old_pre.csv")
write.csv(old_post, file = "cleaned_old_post.csv")
write.csv(new_pre, file = "cleaned_new_pre.csv")
write.csv(new_post, file = "cleaned_new_post.csv")
```

